---
title: "PS270 Final Project"
author: "Ally Jonas"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
```{r}
library(tidyverse)
library(dplyr)
library(kableExtra)
library(nnet)
media_poli_ideology <- read_csv("/users/allyjonas/downloads/SOCIAL MEDIA IMPACT/social_media_political_content_test_data.csv")
```


```{r}
media_poli_ideology
names(media_poli_ideology)
```
## Research Question:
Do the media consumption patterns and previously held notions regarding digital media by global adults impact their political ideologies?

# Introduction
My research analyzes the impact of media consumption and predisposed media beliefs on the political ideologies of adults across the globe. As a mass communications and political science double major, much of the work in my classes involves analyzing and discussing the different ways we can target messages to reach audiences within different political parties. In doing this research, I wanted to observe the statistical impact this communication actually has on the broader public. My hypothesis is that those who prefer more informative political content (e.g. news articles, opinion pieces) will lean more left whereas those who prefer more humorous or entertainment based political content (e.g. memes) will lean more right. Parallel to this, I think those who prefer content centered around the environment and healthcare will lean more left, and those who prefer content regarding immigration and the economy will lean more right. I think this because liberals have statistically shown higher concern for issues such as climate change and lower healthcare costs in the past, where as conservative oftentimes express more concern about their country's economic state.  Additionally, I believe those who trust social media a lot or somewhat will lean left, and those who do not will lean more right. Due to the rise of TruthSocial and other conservative adjacent social-networking platform, I believe conservatives demonstrate an on average larger lack of trust for social media platforms than liberals. Finally, I believe those who are use social media less but interact with political content more frequently on the platforms will be more left leaning, whereas those who are use social media more but interact with political content on it less often will lean more right. I believe this purely on observation. From viewing the people I follow on my social media in comparison to my those my conservative parents do, I notice more liberals posting about current events and political issues on my own feed than conservatives posting political content on my parents'. Although there's a large chance this is simply a generational difference, I still felt it worth investigating. 

# Data

My data set comes from the website Kaggle under the title, “Social Media Political Content Analysis Dataset.” The date in which the survey data was collected and when it was posted to Kaggle is not available on the website. My key dependent variable, political affiliation, comes from self-reported survey data. Additionally, my independent variables, content preference (measured by types of political content and political content topics), trust in social media (measured by trust in social media), and frequency of interaction with political content (measured by social media usage frequency, and political content interaction frequency) were collected as self-reported survey data.

### Dependent Variable Summary Plot

```{r}

party_distrubution_plot <- ggplot(media_poli_ideology, aes(x = `Political Affiliation`)) +
  geom_bar(fill = "#6B9272", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Political Affiliation",
       x = "Political Affiliation",
       y = "Count") +
  theme(
     plot.background = element_rect(fill = "#AD9381", color = NA),
    plot.title = element_text(family = "Georgia", face = "bold", size = 14),
    axis.title.x = element_text(family = "Georgia", face = "bold", size = 12),
    axis.title.y = element_text(family = "Georgia", face = "bold", size = 12),
    axis.text.x = element_text(family = "Helvetica", angle = 45, hjust = 1),
    axis.text.y = element_text(family = "Helvetica"),
      plot.subtitle = element_text(family = "Helvetica")
  )
party_distrubution_plot

party_distrubution_table <- table(media_poli_ideology$`Political Affiliation`)

knitr::kable(party_distrubution_table, caption = "Political Affiliation Table")
```

In this data set, there are 918 respondents who self-identify as conservative, 934 as independent, and 948 as liberal. These findings and this plot show that the categories in my dependent variable are spread approximately evenly, as none received an overly dominant amount of respondents. In turn, this makes my dependent variable well-suited for statistical testing. This plot is cross-sectional, as it visualizes respondents political affiliations at a singular point in time (whenever they took the survey).  

# Results

My primary areas of interest are how certain media consumption patterns are associated with and potentially impact political affiliation in the United States, United Kingdom, Canada, and Australia. With this, my primary relationship of interest is as follows:

## Effect of frequency of social media usage and interaction with political content on political affiliation

```{r}
social_inter_plot <- ggplot(media_poli_ideology, aes(x = `Political Affiliation`)) + 
  geom_bar(aes(fill = `Social Media Usage Frequency`), position = "dodge") +
  facet_wrap(~`Interactions Frequency`, scales = "free_y") + 
  labs(
    title = "Social Media Usage and Interaction Frequency by Political Affiliation",
    x = "Political Affiliation",
    y = "Count",
    fill = "Social Media 
Usage Frequency"
  ) +
  scale_fill_manual(values = c("0-1 hour" = "#6B9272", "1-2 hours" = "#ab560a", 
                              "2-4 hours" = "#8e4d3a", "4+ hours" = "#4e88ab")) +
  theme_minimal() +
  theme(
     plot.background = element_rect(fill = "#AD9381", color = NA),
    plot.title = element_text(face = "bold", size = 10, family = "Georgia"),
    axis.title.x = element_text(size = 7, family = "Georgia"),
    axis.title.y = element_text(size = 8, family = "Georgia"),
    axis.text.x = element_text(face = "bold", size = 6, angle = 45, hjust = 1, family = "Helvetica"),
    axis.text.y = element_text(size = 4), family = "Helvetica",
strip.text = element_text(family = "Helvetica", size = 8), 
legend.title = element_text(family = "Georgia", size = 6),
legend.text = element_text(family = "Helvetica", size = 4), legend.key.size = unit(0.5, "cm"),  
  panel.spacing = unit(1, "lines"),     
  plot.margin = margin(1, -1, 1, 1))


frequency_sum_table <- media_poli_ideology |>
  group_by(`Political Affiliation`) |>
  summarise(
    avg_social_media_usage_0_1hr = mean(`Social Media Usage Frequency` == "0-1 hour", na.rm = TRUE),
    avg_social_media_usage_1_2hr = mean(`Social Media Usage Frequency` == "1-2 hours", na.rm = TRUE),
    avg_social_media_usage_2_4hr = mean(`Social Media Usage Frequency` == "2-4 hours", na.rm = TRUE),
    avg_social_media_usage_4plushr = mean(`Social Media Usage Frequency` == "4+ hours", na.rm = TRUE),
    
    avg_interaction_rarely = mean(`Interactions Frequency` == "Rarely", na.rm = TRUE),
    avg_interaction_few_times = mean(`Interactions Frequency` == "Few times a week", na.rm = TRUE),
    avg_interaction_once_a_day = mean(`Interactions Frequency` == "Once a day", na.rm = TRUE),
    avg_interaction_several_times = mean(`Interactions Frequency` == "Several times a day", na.rm = TRUE)
  )


social_inter_plot
knitr::kable(frequency_sum_table,
             format = "html") |>
  kable_styling("striped", full_width = FALSE) |>
  scroll_box(width = "100%", height = "300px")
```
In the 0-1 hour category, liberals use social media 25.2%, conservatives 24.3%, and independents 23.9% whereas in the 1-2 hour category, liberals use social media 27.1%, conservatives 25.5%, and independents 24.7%. Conservatives and independents use social media 25.9% in the 2-4 hour category, with liberals using it 22.9%. Finally, in the 4+ hour category, independents use social media 25.5%, liberals 24.7%, and conservatives 24.3%.

In regards to political content interaction, conservatives interact with political content rarely 27.1%, liberals 26.8%, and independents 24.6%. Independents interact with political content a few times a week at 28.1%, liberals 26.3%, and conservatives 24.8%. In the once a day category, conservatives interact with political content 26.4%, independents 25.8%, and liberals 22.2%. Lastly, liberals interact with political content 24.8% in the once a day category, with conservatives and independents at 21.7% and 21.5% respectively. 

According to the data, liberals use social media less than independents and conservatives. Additionally, despite their low social media usages, liberals tend to interact with political content on social media more frequently than conservatives or independents. 

## Logical multinomial regression on the social media usage frequency and political content interaction data

```{r}
media_poli_ideology$`Political Affiliation` <- factor(media_poli_ideology$`Political Affiliation`, 
levels = c("Conservative", "Independent", "Liberal"))

levels(media_poli_ideology$`Political Affiliation`)

media_poli_ideology$`Social Media Usage Frequency` <- factor(media_poli_ideology$`Social Media Usage Frequency`)

media_poli_ideology$`Interactions Frequency` <- factor(media_poli_ideology$`Interactions Frequency`)

model <- multinom(`Political Affiliation` ~ `Social Media Usage Frequency` + `Interactions Frequency`, 
                  data = media_poli_ideology)

regression_model <- summary(model)
regression_model
```
In this regression, the coefficients depict the log odds o the likelihood of being a liberal or independent in a category, in comparison to being a conservative. 

The coefficients for social media usage frequency highlight how the frequency of social media usage (relative to the baseline of 0-1 hour a day) impact the likelihood of someone being a liberal or independent.

The coefficients for political content interaction frequency showcase the impact of how often someone interacts with political content on social media (relative to the baseline of "rarely") impacts that likelihood of them being a liberal or independent. 

The intercept for being an independent is -0.096, meaning the log odds of a respondent being an independent is slightly less than those of being a conservative. Since the standard of error for this statistic is 0.123, it is above the  significance level of 0.05 and means the null hypothesis cannot be rejected, making this value not statistically significant. In contrast, the intercept for being a liberal respondent is 0.0053, indicating the log odds of being a liberal respondent as opposed to a conservative is slightly higher. Despite this, the standard error is 0.120, therefore making this value not statistical significant. For liberals, using social media 1-2 hours a day slightly increases the likelihood for being a liberal as opposed to a conservative, as the coefficient is 0.026, although with a 0.130 standard error, this finding is not statistically significant. In contrast, spending 1-2 hours on social media a day as an independent reduces the odds of being an independent in comparison to a conservative by -0.016, but again boasts a not statistically significant standard error of 0.133. If you use social media 2-4 hours a day, you are more likely to be a conservative as opposed to a liberal, as the coefficient is -0.148, but once again is not a statistically significant finding with the standard error being 0.132. In contrast, you are slightly more likely to be an independent than a conservative if you use social media for 2-4 hours a day as the coefficient is 0.016, although having a not statistically significant standard error of 0.132. Those who use social media 4+ hours a day are less likely to be liberals than conservatives with a coefficient of -0.020, with a not statistically significant standard error of 0.132. Those who use social media 4+ hours a day are more likely to be independents than conservatives, with a coefficient of 0.066 despite holding a not statisically significant standard error of 0.133. 

Those who interact with political content a few times a week are more likely to be independent than conservative, with a coefficient of 0.219 and a not statistically significant standard error of 0.129. Similarly, those who interact with political content on social media a few times a week are more likely to be liberal than conservative, with a coefficient of 0.068, despite the not statistically significant standard error of 0.128. Respondents who interact with political social media content once a day are more likely to be independent than conservative, with a coefficient of 0.075 and a not statistically significant standard error of 0.129. Surprisingly, those who interact with political social media content once a day are less likely to be liberal than conservative, with a coefficient of -0.157 although having a not statistically significant standard error of 0.130. Finally, those who interact with political content on social media several times a day are more likely to be independent than conservative, with a coefficient of 0.089 and a not statistically significant standard error of 0.136. Respondents who interact with political content on social media several times a day are also more likely to be liberal than conservative, with coefficient of 0.146 and a not statistically significant standard error of 0.131.

For liberals, social media usage of 2-4 hours a day decreases the overall likelihood of being liberal. In the 1-2 hour category, it slightly increases this likelihood. The 4+ hour category slightly decreases liberals likelihoods' of being a liberal, but by a very small and negligible margin. 

For independents, using social media both 1-2 hours a day and 4+ hours a day slightly increases the likelihood of being an independent, while using social media for 2-4 increases and independents likelihood of being independent by a razor-thin, also negligible margin. 

The frequency of interacting with political content on social media has varied effects on liberals. Interacting either occasionally (a few times a week) or very often (several times a day), increases the likelihood of being a liberal. Despite this, interacting moderately with political content (once a day), decreases the overall likelihood of being a liberal. 

Additionally, the frequency of social media content interactions on increases the likelihood of being an independent if the interactions occur a few times a week or several times a day. Interacting with social media content once still increases the likelihood of bein an independent, but by a much smaller margin than the other two measures in the category. 

My main coefficient of interest is that liberals who interact with social media content once a day as oppose to several times a day or a few times a week, are less likely to be liberal overall. The coeffecient is -0.157, and is of interest to me due to is vast difference from the other two coefficients. The data for liberals interacting with social media content a few times a week and several times a day suggests that they are more likely to be liberal. This could be due to behavioral differences, being exposed to liberal content once a day allows for a mix of other types of content to be interacted with - political or otherwise. This other content could potentially contain conservative messages and inflict conservative viewpoints on the consumers, causing them to align less with the liberal ideology. With this, its then assume that those interacting with liberal content only a few times a week simply don't use social media much, and due to the lack of conservative content they'd be exposed to their viewpoints would not change. Although it wasn't statistically significant, the data did highlight the liberals use social media slightly less than the other affiliations, so this is a fair assumption to make. Finally, those who interact with political content several times a day are likely very politically involved. With this, their algorithms on various social networking platforms are likely recommending them content which aligns with their viewpoints, subsequently not exposing them to other opinions and keep their viewpoints the same. Furthermore, those who are heavily involved in politics likely have strong opinions already, and seeing an occasionally post from a commentator with opposing beliefs would likely not change their stances. 

According to the data overall, it is inconclusive whether social media usage and political content interaction frequency impact political affiliation, as none of the results found within the regression were above the significance level of a = 0.05, therefore meaning we cannot reject the null hypothesis and the result is rendered not statistically significant. Despite this, the plot, the percentage points, and the coefficients found do show slight variations between the political affiliations and there overall social media usage and political content interaction frequency. For liberals, more frequent social media usage and less frequent political content interactions decreases their likelihood of being liberals, wheres more frequent social media use and more frequent interactions increase the likelihood of identifying as an independent. Despite this, the data is not significant enough to support either of my claims: those who use less social media but interact with more political content on the platforms lean left, whereas those who use social media more but interact with political content on it less lean more right. 

None of these coefficient represent a causal effect or should be interpreted causally, as the data being analyzed is observational (not from an experiment). With this, observational date is often a victim of confounding. In the context of my data, for example, many outside influences such as socioeconomic status and income levels could impact my dependent variable of political affiliation. Additionally, logistic regression models assume there is a clear and predictable link between the predictors (in this case social media usage and interaction frequency) and the chances of holding a certain political affiliation. The relationship between these predictors and political affiliation is non-linear however, as many other factors influence this variable. The regression model is not good at capturing these not-linear relationships however - which can lead to misleading data. 

### Although having one primary relationship of interest within this research, their were two other relationships I wanted to examine as well, they are as follows: 


## Political affiliation and social media content type and topic preferences

```{r}
content_type_topic_avg_table <- media_poli_ideology |>
  group_by(`Political Affiliation`) |>
  summarise(
    avg_enviro_interest = mean(`Political Content Topics` == "Environment", na.rm = TRUE),
    avg_healthcare_interest = mean(`Political Content Topics` == "Healthcare", na.rm = TRUE),
    avg_immigration_interest = mean(`Political Content Topics` == "Immigration", na.rm = TRUE),
    avg_economy_interest = mean(`Political Content Topics` == "Economy", na.rm = TRUE),
    avg_opinion_pieces = mean(`Types of Political Content` == "Opinion pieces", na.rm = TRUE),
    avg_memes = mean(`Types of Political Content` == "Memes", na.rm = TRUE),
    avg_news_articles = mean(`Types of Political Content` == "News articles", na.rm = TRUE)
  )
knitr::kable(content_type_topic_avg_table, caption = "Average Interest by Political Affiliation 
for Political Topics and Content Types", format = "html") |>
  kable_styling("striped", full_width = FALSE) |>
  scroll_box(width = "100%", height = "300px")

party_distrubution_plot <- ggplot(media_poli_ideology, aes(x = `Types of Political Content`, fill = `Political Affiliation`)) +
  geom_bar(position = "dodge") +
  facet_wrap(~`Political Content Topics`) + 
  labs(
    title = "Political Affiliation by Content Type and Political Topic",
    x = "Type of Political Content",
    y = "Count",
    fill = "Political Affiliation"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Conservative" = "#A65E5E", "Liberal" = "#5E99A6", "Independent" = "#6A547B")) +
  theme(
     plot.background = element_rect(fill = "#AD9381", color = NA),
    plot.title = element_text(face = "bold", size = 12, family = "Georgia"),
    axis.title.x = element_text(size = 10, family = "Georgia"),
    axis.title.y = element_text(size = 10, family = "Georgia"),
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1, family = "Helvetica"),
    axis.text.y = element_text(size = 8), family = "Helvetica",
strip.text = element_text(family = "Helvetica"), 
legend.title = element_text(family = "Georgia"),
legend.text = element_text(family = "Helvetica")
)


party_distrubution_plot
knitr::kable(content_type_topic_avg_table,
             format = "html") |>
  kable_styling("striped", full_width = FALSE) |>
  scroll_box(width = "100%", height = "300px")
```
According to this plot and this table placing the plot data into a numerical format, conservatives are interested 25.93%, independents 23.23%, and liberals 25.21% in environmental political content. In terms of content regarding healthcare topics, conservatives are interested 23.09%, independents 24.09%, and liberals 23.84%. With immigration, conservatives are interested 26.69%, independents 26.65%, and liberals 26.48% in this content topic. Finally, in terms of the economy, conservatives are interested  24.29%, independents 26.02%, and liberals 24.47%.


In terms of political content preferences, there was slightly more variation. In terms of opinion pieces, conservatives preferred this type of content at a rate of 30.50%, independents 31.16%, and liberals 33.23%. When it comes to memes, conservatives preferred them by 33.22%, independents 32.33%, and liberals 32.59%. Finally, conservatives preferred news articles by 36.27%, independents 36.51%, and liberals 34.18%. If we add the averages of both informative political content categories together, opinion pieces and news articles, from each affiliation, the preference percentages are conservatives 66.77%, independents 67.67%, and liberals 67.41%.

Regarding the plot, it shows the spread of political content topics respondents consume most, and the preferred type of content in which they consume these topics. Overall, liberals, conservatives, and independent prefer news articles over other types of political content. Despite this, conservatives' second most preferred content type was memes, and they preferred this genre more than the other two affiliations. Liberals were the most likely to prefer opinion pieces, as this was their second most preferred type of political content, with memes being the lowest. Finally, independents content consumption patterns reflect those of conservatives, albeit to a lesser extent. Their second most preferred type of political content was memes, with their least preferred being opinion pieces. 

In terms of topics, all affiliations were most interest in  immigration content. Of the affiliations, conservatives were the most likely to prefer content relating to immigration, while this number further spread across other categories for liberals and independents. Liberals and conservatives showed interest in environmental content the second most, with conservatives also preferring this category the most in comparison to the other affiliations. Independents preferred economic content the second most, and were the most likely to consume this type of content across all affiliations. Both liberals and conservatives were least likely to consume content relating to healthcare, whereas independents were least likely to consume content relating to the environment.

### Analysis of variance test for political content topics

```{r}
enviro_anova <- aov(`Political Content Topics` == "Environment" ~ `Political Affiliation`, data = media_poli_ideology)
 

healthcare_anova <- aov(`Political Content Topics` == "Healthcare" ~ `Political Affiliation`, data = media_poli_ideology)

immigration_anova <- aov(`Political Content Topics` == "Immigration" ~ `Political Affiliation`, data = media_poli_ideology)

economy_anova <- aov(`Political Content Topics` == "Economy" ~ `Political Affiliation`, data = media_poli_ideology)

anova_topic_results <- data.frame(
  Topic = c("Environment", "Healthcare", "Immigration", "Economy"),
 P_value = c(
    summary(enviro_anova)[[1]][["Pr(>F)"]][1],
    summary(healthcare_anova)[[1]][["Pr(>F)"]][1],
    summary(immigration_anova)[[1]][["Pr(>F)"]][1],
    summary(economy_anova)[[1]][["Pr(>F)"]][1])) 
knitr::kable(anova_topic_results)
```
After running an Analysis of Variance test, the p-value for environmental content preference is ~0.379, healthcare ~0.872, immigration ~0.994, and the economy ~0.639. All of these values are above the significance level of a = 0.05, meaning we cannot reject the null hypothesis and these findings are not statistically significant. None of the evidence supports my hypothesis that those who prefer healthcare and environmental content lean more left, and those who prefer immigration and economic content lean more right.

### Analysis of variance test for political content types
```{r}
opinion_anova <- aov(`Types of Political Content` == "Opinion pieces" ~ `Political Affiliation`, data = media_poli_ideology)


memes_anova <- aov(`Types of Political Content` == "Memes" ~ `Political Affiliation`, data = media_poli_ideology)

news_articles_anova <- aov(`Types of Political Content` == "News articles" ~ `Political Affiliation`, data = media_poli_ideology)

anova_type_results <- data.frame(
  Type = c("Opinion Pieces", "Memes", "News Articles"),
  P_value = c(summary(opinion_anova)[[1]][["Pr(>F)"]][1],
    summary(memes_anova)[[1]][["Pr(>F)"]][1],
    summary(news_articles_anova)[[1]][["Pr(>F)"]][1]))
knitr::kable(anova_type_results)
```

After running an analysis of variance test on the data regarding political content type preferences, the p-values were as follows: opinion pieces ~0.416, memes ~0.916, news articles ~0.509. The data suggests that their is no statistically significant evidence to support my claim that those who prefer entertaining political content, such as memes, tend to lean more right. The p-values for all measurable components are above the significance level of a = 0.05, meaning we cannot reject the null hypothesis and our results are not statistically significant. None of the data analyzed supports my hypothesis that those who prefer informative political content (e.g. opinion pieces and news articles) lean more left, whereas those who prefer entertaining political content (e.g. memes) lean more right. In addition to the p-values reflecting this, independents and conservatives prefer informative content at either comparable or slightly higher rates than liberals after adding the percentage point of both categories within each affiliation, further rejecting my claims. 


## Effect of social media trust levels on political affiliation

```{r}
poli_trust_plot <- ggplot(media_poli_ideology, aes(x = `Political Affiliation`, fill = `Trust in Social Media`)) +
  geom_bar(position = "dodge") +  # Position bars side by side
  labs(
    title = "Trust in Social Media by Political Affiliation",
    x = "Political Affiliation",
    y = "Count",
    fill = "Trust in Social Media"
  ) +
  scale_fill_manual(values = c("Trust a lot" = "#6B9272", "Trust somewhat" = "#ab560a", "Do not trust" = "#8e4d3a")) +
  theme_minimal() +
  theme(
   plot.background = element_rect(fill = "#AD9381", color = NA),
    plot.title = element_text(face = "bold", size = 12, family = "Georgia"),
    axis.title.x = element_text(size = 10, family = "Georgia"),
    axis.title.y = element_text(size = 10, family = "Georgia"),
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1, family = "Helvetica"),
    axis.text.y = element_text(size = 8), family = "Helvetica",
strip.text = element_text(family = "Helvetica"), 
legend.title = element_text(family = "Georgia"),
legend.text = element_text(family = "Helvetica")
)

media_poli_ideology$`Trust in Social Media numeric` <- 
  as.numeric(factor(media_poli_ideology$`Trust in Social Media`, 
                    levels = c("Do not trust", "Trust somewhat", "Trust a lot"), 
                    labels = c(1, 2, 3)))

trust_social_averages_table <- media_poli_ideology |>
  group_by(`Political Affiliation`) |>
  summarise(
    avg_trust_in_social_media = mean(`Trust in Social Media numeric`, na.rm = TRUE)
  )

poli_trust_plot
knitr::kable(trust_social_averages_table)
```
Across all three affiliations, most people trust social media platforms a lot, with independents being the mostly likely out of the three groups to trust social media information. After converting this data to a numerical format (Do not trust = 1, Trust somewhat = 2, Trust a lot = 3), the average trust levels for each political affiliation were as follows: conservatives = 2.032680, independents = 2.035332, and liberals 2.021097.

Independents and liberals were around the same in the number of respondents claiming they did not trust social media, but liberals were the least likely to trust social media platforms as a whole on average. This is likely because independents were the group least likely to trust social media somewhat, while liberals were the most. Conservatives trust social media slightly less than independents on average, and were more likely to trust social media somewhat then them. However, conservatives fall behind independents by a small margin, and were the group second most likely to trust information from social media. Among respondents, conservative identifying individuals were the least likely to report not trusting social media. 

### Analysis of variance in social media trust levels

```{r}
anova_social_results <- aov(`Trust in Social Media numeric` ~ `Political Affiliation`, data = media_poli_ideology)


social_anova_results <- data.frame(
  "Test" = c("Trust in Social Media 
             by Political Affiliation"),
   P_value = c(summary(anova_social_results)[[1]][["Pr(>F)"]][1]
   ))

knitr::kable(social_anova_results)
```

Despite my prior statements, in terms of my hypothesis, trust in social media does not appear strongly correlate with political affiliation. This is because the average trust in social media across all three political affiliations is very similar, and my findings are not statistically significant enough to suggest that people who trust social media more lean left, and those who don't lean right. The findings aren't considered statistically significant because in running an Analysis of Variance on the data, the p-value come out to ~0.923. This is largely above the significance level threshold of a = 0.05, and means there is not enough evidence to reject the null hypothesis - therefore rendering the results not statistically significant.

## Conclusion 

In my hypothesis, I stated I though people would lean further left if they used social media less often but interacted more with political content on the platforms when they were used, preferred to get their political content from the more informative types available (e.g. news articles and opinion pieces) and prefered content relating to healthcare or the environment, and had higher levels of trust in information from social media. I thought people would lean further right if they used social media more but interacted with political content less on the platforms, preferred more entertaining means of consuming political content (e.g. meme) and was interested in this content if it centered immigration or the economy, and had lower levels of trust in social media. 

My findings were inconclusive as to whether those who use social media less but interacted with political content more on the platforms leaned further left and those who used social media more but interacted with political content less on the platforms leaned more right. None of my data for this was statistically significant, and therefore I cannot draw any concrete conclusions from my results. 

In addition to this, my findings as to whether those who preferred more informative types of political content (e.g. new articles and opinion pieces) and topics relating to the environment or healthcare lean more left, and those who preferring entertain political content (e.g. memes) and topics surrounding immigration and the economy lean more right were also inconclusive. None of the data from this section was statistically significant either, and I was again cannot make any concrete judgement. Furthermore, certain data collected may disprove my original claims, as independents and conservatives prefer informative political content at comparable and/or higher rates than liberals. Additionally, all political affiliations were most interested in topics relating to immigration, further disproving my hypothesis. 

Finally, my findings as to whether those who have higher trust for social media lean further left, and those who have lower trust for it lean right was also inconclusive. Echoing the previous two sections, my findings were not statistically significant and therefore I cannot draw any official conclusions from my data. Additionally, liberals were the group least likely to trust social media as a whole, once again further disproving my original claim.

Since my data was an observational study, and the relationships between my predictors and dependent variable were non-linear, I'm confident my data was largely confounded. As I stated previously, political affiliation is formed based on a number of different factors (socioeconomic status, familial background, employment status and/or field, etc) and is therefore difficult to measure in an observational study where the kinds of people taking this test is unknown. The threat to interference with this is that if the confounders are not controlled, it leads to misleading data. Obviously my confonders were not able to be controlled, and much of my data is subsequently, most-likely misleading. Additionally, simple phrases/numerical scale measurements on factors like social media trust, political content preferences, and political content interaction frequency, are not able to fully convey the nuances and complexities behind these behaviors. One exmaple of this is that political content preferences may vary across topics (for instance: I'm not going to seek out an informative piece picking apart "Kamala is brat"), or social media trust levels may change platform to platform. With this, the biggest threat to interference is that if measurements aren't precise enough, they can easily miss important behaviors in the relationships being studied, causing the result to inaccurately reflect these relationships. 

If I had more time and money, I'd put together my own group to conduct this research on to mitigate the potential for confounding. I'd have them take the survey, then have a group adjust their media consumption habits for an extended period of time, and finally have them come back later and take the survey again. This would help me accurately determine the effects of media consumption on political ideology, as it would be measured as a treatment effect over a longer period of time versus an observational study measured once with no indication of the people who you've received data from. 

