---
title: "PS270 Final Project"
author: "Ally Jonas"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
## Research Question:
Does the type of content US adults view on social media impact their political ideologies?

## Hypothesis:
With the growing amount of adults in the United States who use and consume important information from social media, I think that those who prefer informative or political content on social media will lean more Democratic, and those who prefer memes and more humorous content will lean more Republican. I feel this way because on average, Democrat voter have more extensive educational background than Republicans, which leads me to believe they would seek out education content on a more frequent basis. These results were measured through survey data posted to Kabble under the name Social Media Political Content Analysis Dataset. 

## Explanation:
I will be examining adults over 18, their political affiliations (Liberal, Conservative, or Independent), their preferred content on social media (New articles, opinion pieces, memes), and the frequency at which they use social media. Once again, all of this information was gathered from survey data posted on Kabble under the name Social Media Political Content Analysis Dataset. 

## Outcome Variable:
My outcome variable is the party ideologies of those who prefer news articles, opinion pieces, and memes on social media. This variable, like my others, was measured through survey data posted to Kabble under the name Social Media Political Content Analysis Dataset.

## Patterns:
A pattern that would support my hypothesis is that the majority of people with a degree beyond a bachlors (masters and PhD) are reportedly Liberal. A pattern that may harm my hypothesis is that most people Liberals say the news they’re seeing on social media isn’t very accurate, which would lead me to assume it did not impact their political beliefs. 

```{r}
library(tidyverse)
library(dplyr)
media_poli_ideology <- read_csv("/users/allyjonas/downloads/SOCIAL MEDIA IMPACT/social_media_political_content_test_data.csv")
```


```{r}
media_poli_ideology
names(media_poli_ideology)
```
## Poltitical Ideology by Preferred Social Network Bar Chart
```{r}
ggplot(media_poli_ideology, aes(
  x = `Political Affiliation`,
  fill = `Political Affiliation`
)) + geom_bar() + facet_wrap(~ `Preferred Social Media`, scales = "free_y") +
  labs(
    title = "Political Ideology by Preferred Social Network",
    x = "Preferred Social Network",
    y = "Count",
    fill = "Poltical Affiliation") + theme(
      axis.text = element_text(size = 3),
      plot.title = element_text(size = 10),
      axis.title = element_text(size = 7))
```
## Recency of Political Content Exposure by Political Affiliation

```{r}
media_poli_ideology <- media_poli_ideology |>
   mutate(`Recency of Exposure` = case_when(
    `Recency of Exposure` == "Within the last hour" ~ 1,
    `Recency of Exposure` == "Within the last 24 hours" ~ 2,
    `Recency of Exposure` == "Within the last week" ~ 3,
    `Recency of Exposure` == "Longer than a week ago" ~ 4,
    TRUE ~ NA_real_ ))

line_data <- media_poli_ideology |>
  group_by(`Political Affiliation`) |>
  summarize(mean_recency = mean(`Recency of Exposure`, na.rm = TRUE), .groups = "drop")

recent_poli_content <- ggplot(line_data, aes(
    x = `Political Affiliation`, 
    y = mean_recency, 
    group = 1, 
    color = `Political Affiliation`)) +
  geom_line(size = 1.2) + geom_point(size = 3) +  
  labs(
    title = "Average Recency of Political Content Exposure by Political Affiliation",
    x = "Political Affiliation",
    y = "Average Recency of Exposure"
  ) +
  theme_minimal()
recent_poli_content
```

This line plot depicts the recency of political content exposure on average by political party. If people had engaged with political content within the hour they were given a 1, within the past 24 hours a 2, within the past week a 3 and longer than a week ago a 4. Conservatives had the highest average with a 2.52, with independents at 2.47 and liberals leading by a small margin at 2.46. People who identify as liberal are more likely to have viewed political content more recently than people who identify as both independent or conservatives. Furthermore, conservatives were the least likely to have viewed political content recently. These findings highlight the differences in media consumption patterns between political groups, and their potential implications on party affiliation. 
